\documentclass[times,10pt]{sigplanconf}

\input{preamble.tex}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\doi{nnnnnnn.nnnnnnn}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{\name}                        % 'preprint' option specified.

\title{System \name: A Simple Core Language for Extensibility}
%%\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Name2\and Name3}
           {Affiliation2/3}
           {Email2/3}

\maketitle

\begin{abstract}
  
  Over the years there have been various proposals for \emph{design
    patterns} to improve \emph{extensibility} of programs.
  Examples include \emph{Object Algebras}, \emph{Modular Visitors} or
  Torgersen's design patterns using generics.
  Although those design patterns give practical
  benefits in terms of extensibility, they also expose limitations in
  existing mainstream OOP languages. Some pressing
  limitations are: 1) lack of good mechanisms for
  \emph{object-level} composition; 2) \emph{conflation of 
    (type) inheritance with subtyping}; 3) \emph{heavy reliance on generics}.

  This paper presents System \name: an extension of System F with
  \emph{intersection types} and a \emph{merge operator}.  The goal of System \name
  is to study the minimal language constructs needed to support
  various extensible designs, while at the same time addressing the
  limitations of existing OOP languages. To address the lack of good
  object-level composition mechanisms, System \name uses the merge
  operator to do dynamic composition of values/objects. Moreover, in
  System \name type inheritance is independent of subtyping, and an
  extension can be a supertype of a base object type.  Finally, System
  \name replaces many uses of generics by intersection types or
  conventional subtyping. System \name is formalized and
  implemented. Moreover the paper shows how various extensible designs
  can be encoded in System \name.

\end{abstract}

%%\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore,
% you may leave them out
%%\terms
%%Design, Languages, Theory

%%\keywords
%%Intersecion Types, Polymorphism, Type System

\input{sections/introduction.tex}
\input{sections/overview.tex}
\input{sections/application.tex}
\input{sections/fi.tex}
\input{sections/translation.tex}
\input{sections/implementation.tex}
\input{sections/related-work.tex}
\input{sections/conclusion.tex}

\input{sections/bibliography.tex}

\begin{comment}
\acks

Acknowledgments, if needed.
\end{comment}

\clearpage
\onecolumn
\appendix
\input{sections/type-wellformedness.tex}
\input{sections/target-type-system.tex}
\input{sections/proofs.tex}

\end{document}
